Co-citations in context

### 2019-06-08 cocit_r.tex
revisions:

% ie,
%\correspondingauthor{Ritwik K. Niyogi}{ritwik.niyogi@gatsby.ucl.ac.uk}

@@

\citep{uzzi_atypical_2013} used a novel approach for co-citation analysis using, as input, 17.9 million articles and their cited references, from the Web of Science (WoS)  to characterize a subset of highly cited articles with respect to both novel and conventional combinations of prior research. In this study, co-citation frequencies of references in were calculated and represented as journal pairs (observed co-citation frequencies). Expected values of these observed frequencies were then computed from Monte Carlo simulations under a random graph model. Observed frequencies were then normalized (shifted and scaled) to expected values and termed \emph{z-scores} (Materials and Methods). Thus, every article was associated with multiple z-scores corresponding co-cited journal pairs in its citations. For each article, positional statistics of these z-scores were calculated to describe conventionality: high conventionality (HC) if the median z-score for an article was greater than the median of median z-scores of all articles and low conventionality for the converse (LC). Similarly,  an article was labeled high novelty (HN) if the tenth percentile of its z-scores  was negative and low novelty (LN) for the converse. Accordingly, each article was annotated with respect to conventionality and novelty, e.g, HCHN (denoting that the article exhibits high conventionality and high novelty), with all four combinations being possible. 
\cite{uzzi_atypical_2013} observed that HCHN articles were twice as likely to be highly cited, suggesting that novel combinations of ideas flavoring a body of conventional thought were `nearly universal' recipe for impact.

Key to Uzzi {\em et al.}, however, is their random graph model and its underlying assumptions. The citation switching algorithm used to generate a null model in Uzzi  {\em  et al.} elegantly preserves the number of publications, the number of references in each publication, the year of publication of both publications and references, and the disciplinary composition of the references cited in these publications. Random substitution of the references used to generate expected values is equiprobable with respect to disciplinary origin and citation count.  The model would permit substitution of a reference in quantum physics with equal probability by a reference in the related quantum chemistry field or in some entirely different field, such as classical literature, evolutionary biology, or anthropology. Such substitutions poorly model documented the disciplinary nature of scientific endeavor and citation behavior~\citep{wallace_lariviere_gingras_2012,moed_measuring_2010,klavans_research_2017,garfield_1979}. In addition, under this random model, a reference cited over 100 times in a given year is selected with the same probability as a reference cited only once, which is also inconsistent with citation behavior and inconsiderate of the power law or lognormal citation distributions described in the literature~\citep{stringer_statistical_2010,perline_strong_2005}.  Accordingly, the expected value calculations generated by the \cite{uzzi_atypical_2013} simulations and used in characterizing articles in terms of conventionality and novelty can be reasonably questioned on grounds of model mis-specification when subsets of the WoS are analyzed in the context of expected values generated across disciplines.

@@

Despite different methods used to generate expected values, both \citep{uzzi_atypical_2013} and \citep{boyack_vs_uzzi_2014} measured observed co-citation frequencies across the scientific literature without disciplinary constraints and subsequently used normalized frequencies to examine disciplinary subsets. In extending their prior work, we hypothesized that analysis of datasets focused in disciplinary areas would reduce model mis-specification since the pool of citations would be relevant to those disciplinary areas rather than being drawn from all disciplines.

Accordingly, we used keyword searches of the scientific literature to create three citation networks themed around academic disciplines. Within these disciplinary frameworks, we calculated observed and expected co-citation frequencies using a refined random graph model and an efficient Monte Carlo simulation algorithm. Our analysis of these semantically-themed citation networks challenges the assertion of universality, particularly that articles in the HCHN category have the highest probability of being hits. We find, for example, while HC remains highly correlated with hit articles in the immunology and metabolism datasets we constructed, HN is not. In addition, while HN is highly correlated with hit articles in applied physics, HC is not.  Furthermore, we found that the categories demonstrating the highest percentage of hits are not robust with respect to varying parameters that define hit articles and the threshold for highly novel citation patterns.

@@

\section{Materials and Methods}
\emph{Bibliographic data} We have previously developed ERNIE, an open source knowledge platform into which we parse the Web of Science (WoS) Core Collection~\citep{Keserci371955}. WoS data stored in ERNIE spans the period 1900-2019 and consists of over 72 million publications. For this study, we generated an analytical dataset from years 1985 to 2005 from ERNIE. The total number of publications in this dataset was just over 25 million publications (25,134,073), which were then stratified by year of publication. For each of these years, we further restricted analysis to publications of type Article. Since WoS data also contains incomplete references or references that point at other indexes, we also considered only those references for which there were complete records~(Table \ref{tab:summary_data}). For example, WoS data for year 2005 contained 1,753,174 publications, which after restricting to type Article and considering only those references described above resulted in 916,573 publications, 6,095,594 unique references (set of references), and 17,167,347 total references (multiset of references). Given consistent trends in the data, we analyzed the two boundary years (1985 and 2005) and the mid-point (1995) performing 1,000 simulations for each dataset. We also calculated the number of times each of these articles was cited in the first 8 years since publication.

% latex table generated in R 3.5.3 by xtable 1.8-4 package
% Fri May 17 14:28:31 2019
\begin{table}[ht]
\caption{Summary of WoS Analytical Dataset.The number of publications of type Article (UP), unique references (UR), total references (TR) and the ratio of total references to unique references increases monotonically with each year indicating that both the number of documents and citation activity increase over time. Data for reference years is shown in boldface.} 
\label{tab:summary_data}
\centering
\begin{tabular}{|rrrrr|}
  \hline
Year & UP & UR & TR & TR/UR \\ 
  \hline
\textbf{1985} & \textbf{418495} & \textbf{2281297} & \textbf{5615496} & \textbf{2.46} \\ 
\hline
1986 & 402309 & 2316451 & 5708796 & 2.46 \\ 
1987 & 412936 & 2427347 & 5998513 & 2.47 \\ 
1988 & 426001 & 2545647 & 6354917 & 2.50 \\ 
1989 & 443144 & 2673092 & 6749319 & 2.52 \\ 
1990 & 458768 & 2827517 & 7209413 & 2.55 \\ 
1991 & 477712 & 2977784 & 7729776 & 2.60 \\ 
1992 & 492181 & 3134109 & 8188940 & 2.61 \\ 
1993 & 504488 & 3278102 & 8676583 & 2.65 \\ 
1994 & 523660 & 3458072 & 9255748 & 2.68 \\ 
 \hline
\textbf{1995} & \textbf{559685} & \textbf{3692575} & \textbf{9897946} & \textbf{2.68} \\ 
\hline
1996 & 663110 & 4144581 & 11641286 & 2.81 \\ 
1997 & 677077 & 4340733 & 12135104 & 2.80 \\ 
1998 & 693531 & 4573584 & 12728629 & 2.78 \\ 
1999 & 709827 & 4784024 & 13280828 & 2.78 \\ 
 2000 & 721926 & 5008842 & 13810746 & 2.76 \\ 
 2001 & 727816 & 5203078 & 14261189 & 2.74 \\ 
 2002 & 747287 & 5464045 & 15001390 & 2.75 \\ 
 2003 & 786284 & 5773756 & 16024652 & 2.78 \\ 
 2004 & 826834 & 6095594 & 17167347 & 2.82 \\ 
 \hline
\textbf{2005} & \textbf{916573} & \textbf{6629595} & \textbf{19066249} & \textbf{2.88} \\ 
 \hline
\end{tabular}
\end{table}

@@

\emph{Monte Carlo simulations and normalization of observed frequencies}. Analysis was performed on publications from 1985, 1995, and 2005. Building upon prior work~\cite{uzzi_atypical_2013}, all ${n \choose 2}$ reference pairs were generated for each publication, where $n$ is the number of cited references in the publication. These reference pairs were then mapped to the journals they were published in using ISSN numbers to create journal pairs. Where multiple ISSN numbers exist for a journal, the most frequently used one in the WoS was assigned to the journal. In addition, publications containing fewer than two references were discarded. Journal pair frequencies were summed across the dataset to create observed frequencies $(F_{obs})$. We developed a performant citation switching algorithm, \emph{runtime enhanced permuting citation switcher (repcs)} that randomly permuted citations grouped by year of publication to switch citations while preserving the year of publication for both articles and references, the number of publications and the number of references in each dataset and the disciplinary composition of the references in each dataset. The code for the citation switching algorithm~\cite{uzzi_atypical_2013} kindly provided by the authors and referred to as \emph{umsj}~\cite{boyack_vs_uzzi_2014}, was used as a benchmark for the repcs algorithm. Comparative analysis showed that 10 rounds of citation switching in a dataset of 391,860 publications from 1985 completed in 2,186 hours using the umsj algorithm and in less than one hour using the repcs algorithm. In that test, using either the repcs or umsj algorithms resulted in generating simulated values for 75\% of the observed journal pair frequencies. Subsequent rounds of development and testing resulted in 99\% journal pair coverage with 1,000 simulations so we chose use the repcs algorithm in our study and generated 1,000 null models, in contrast to the 10 null models in the previous study, for each dataset we analyzed.

Expected values $(F_{exp})$ for journal pairs were generated by averaging the result of 1,000 simulations and z-scores were then calculated for each journal-pair using the formula $(F_{obs} - F_{exp})/\sigma$ where $\sigma$ is the standard deviation of the frequencies generated by simulation. As a result of these calculations, each publication becomes associated with a set of z-scores corresponding to the journal pairs derived from pairwise combinations of its cited references and positional statistics (quantiles) of z-scores were calculated for each publication. Publications were subsequently labeled according to conventionality and novelty: (i) HC if the median z-score exceeded the median of median z-scores for all publications and LC if the median z-score was equal to or less than the median of median z-scores for all publications, 
and (ii) HN if the tenth percentile of z-scores for a publication was less than zero, and LN if the tenth percentile of z-scores for a publication was greater than zero.  We also investigated multiple definitions of hit articles with hits defined as the 1\%, 2\%, 5\%, and 10\% top-cited articles.  Also, mention different thresholds, 1st and 10th, for novelty threshold.

@@

\subsection{Model Misspecification and the Attributes of Disciplinary Context} We have invoked model mis-specification arising from disciplinarily irrelevant references as a concern. To address this issue, we analyzed the disciplinary composition of the references in our three test datasets, `applied physics', `immunology', and `metabolism' for the years 1985, 1995, and 2005 and observed that 8.4.2\% of the references in the applied physics dataset were classified as belonging to the WoS physical sciences research area. Similarly, 93.7\% and 92.1\% of the references in the immunology and metabolism datasets belonged to the life sciences and biomedicine research area, indicating strongly that references are more likely belong to the same research area as the publications that cite them.

We  also analyzed the consequences of citation shuffling within a disciplinary set or all of the Web of Science.  References in publications belonging to these three datasets were classified using the WoS Extended Subjects classification of 153 subject areas and summarized as a frequency distribution. A single shuffle of the references in these datasets or all the references in the corresponding WoS year slice was performed, using either the rpcs or umsj algorithms, and frequencies computed again. The fold change in subject frequencies of references before and after shuffling was calculated for all 153 Subject categories and summarized as boxplots in \ref{fig:be}. The data show consistently that disciplinary composition of references is preserved when citation shuffling is conducted within a disciplinary network.

We then estimated model mis-specification by calculating the Kullback-Leibler (K-L) divergence between the simulated and observed journal pair frequencies for disciplinary datasets using again, either the references in the disciplinary datasets or those in all of WoS for the simulations~\ref{tab:kld} and observed that K-L divergences when using the WoS background was roughly double compared to the disciplinary network.

We addressed this consideration by analyzing disciplinary subsets of the scientific literature, thereby restricting random selection of references to only those references in the disciplinary network being studied.
What about z-scores? 
We have conjectured that the approach of Uzzi et al. for generating journal-pair z-scores is misspecified in its sampling from a broad dataset and its disregard for the frequency with which journal pairs are citedOur principal consideration was to restrict model misspecification arising from disciplinarily irrelevant references. We addressed this consideration by analyzing disciplinary subsets of the scientific literature, thereby restricting random selection of references to only those references in the disciplinary network being studied.As observed in the Introduction, Monte Carlo simulations that use references from all publications do not account for observed citation practice (sentence needs to be made more eloquent)...

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=0.8\linewidth]{background-effect}     
\caption{\ Intra-network citation shuffling preserves the disciplinary composition of references. Publications of type Article belonging to the three disciplinary networks (i) Applied Physics (ii) Immunology (iii) Metabolism were subject to a single shuffle of all their cited references using either the cited references in these networks as a source of random substitutions (bg\_local) or references from all articles in WoS (bg\_allWoS). Citation shuffling was performed using either our algorithm (repcs) or that of Uzzi (umsj). The disciplinary composition of cited references before and after shuffling was measured as frequencies for each of 153 sub-disciplines (from the extended subject classification in WoS) and expressed as fold difference between citation counts grouped by subject for original (o) and shuffled (s) references using the formula (fold\_difference = $ifelse(o > s, o/s, s/o)$) and rounded to the nearest integer. Null values were set to 1. A fold difference of 1 indicates that citation shuffling did not alter disciplinary composition. Data are shown for articles published in 1985. All four boxplots are generated from 151 observations. Min \& Max values for fold differences were as follows. (i) Arts \& Humanities. bg\_local (1,2), bg\_allWoS (1, 488) (ii) Life Sciences \& Biomedicine. bg\_local(1,1), bg\_allWos (1, 352) Note y-axis: $log_2$ scale.} 
\label{fig:be}
\end{figure}

Figure \ref{fig:scatter} shows that the z-scores for the same journal pair can be positive (negative) when computed with respect to one data set but be negative (positive) for another data set. The journal-pair z-scores in Figure \ref{fig:scatter} have consistent signs for both Immunology and WoS data sets in 71.4\% of the instances and different signs for 28.6\% of the journal pairs. When a journal-pair z-score is negative with respect to one dataset and positive with respect to another dataset, articles citing that pair are more likely to be deemed novel in the first instance and less likely to be deemed novel in the second instance.  It seems that any journal pair should either be indicative of novelty or not, but it should never have contradictory implications that depend on the reference dataset. We view, therefore, these contradictions as symptomatic of inappropriately sampling from a broad dataset and ignoring observed citation patterns.  Figure \ref{fig:scatter} reflects that the WoS data set has approximately 44,000 fewer negative z-scores than does the immunology data set, which contributes to its significantly lower percentage of high-novelty articles.  \emph{The plots I made of the cumulative z-score distributions for Immunology and WoS also could be used to demonstrate this phenomenon, although the scatter plot does this more clearly, and it is more striking.  The scatter plot also encodes more data, that is, the matched z-scores for each journal pair.}

>>>>

% ie,
\correspondingauthor{George Chacko}{netelabs@nete.com}

@@

\citep{uzzi_atypical_2013} used a novel approach for co-citation analysis using, as input, 17.9 million articles and their cited references, from the Web of Science (WoS)  to characterize a subset of highly cited articles with respect to both novel and conventional combinations of prior research. In this study, co-citation frequencies of references in were calculated and represented as journal pairs (observed co-citation frequencies). Expected values of these frequencies were generated from Monte Carlo simulations under a random graph model. Observed frequencies were then normalized (shifted and scaled) to averaged expected values from ten simulations and termed \emph{z-scores}, which were interpreted as suggesting conventional thought if positive and novel thought if negative. Consequently, every article was associated with multiple z-scores corresponding to co-cited journal pairs in its refernces. For each article, positional statistics of z-scores were calculated to set thresholds for a binary (high/low) classification of conventionality (C) using the median and novelty (N) using the tenth percentile of z-scores within an article. Thus, HCLN would denote high conventionality and low novelty with all four combinations being possible. 
\cite{uzzi_atypical_2013} observed that HCHN articles were twice as likely to be highly cited, suggesting that novel combinations of ideas flavoring a body of conventional thought were a 'nearly universal` recipe for impact.

Key to Uzzi {\em et al.}, is their random graph model and its underlying assumptions. The citation switching algorithm used to generate a null model elegantly preserves the number of publications, the number of references in each publication, the year of publication of both publications and references, and the disciplinary composition of the references cited in these publications. However, random substitution of the references used to generate expected values is equiprobable with respect to disciplinary origin and citation count. The model permits substitution of a reference in quantum physics with equal probability by a reference in the related quantum chemistry field or in some entirely different field, such as classical literature, evolutionary biology, or anthropology. Such substitutions poorly model the disciplinary nature of scientific endeavor and citation behavior~\citep{wallace_lariviere_gingras_2012,moed_measuring_2010,klavans_research_2017,garfield_1979}. In addition, under this random model, a reference cited over 100 times in a given year is selected with the same probability as a reference cited only once, which appears inconsistent with the power law or lognormal citation distributions described in the literature~\citep{stringer_statistical_2010,perline_strong_2005}.  Accordingly, the expected value calculations used to characterize articles in terms of conventionality and novelty can be reasonably questioned on grounds of model mis-specification when disciplinary subsets of the WoS are analyzed in the context of expected values generated across all of WoS.

@@

Despite different methods used to generate expected values, both of these key preceding studies (vide supra) measured co-citation frequencies across the scientific literature without disciplinary constraints and subsequently analyzed disciplinary subsets. In extending their prior work, we hypothesized that normalization using expected values drawn from datasets focused in disciplinary areas, rather than all of WoS, would reduce model mis-specification.  Accordingly, we used keyword searches of the scientific literature to create three citation networks themed around academic disciplines. Within these disciplinary frameworks, we calculated observed and expected co-citation frequencies using a refined random graph model and an efficient Monte Carlo simulation algorithm. Our analysis of these semantically-themed citation networks challenges the assertion of universality. Under our conditions of reduced model misspecification, we observe that while HC remains highly correlated with hit articles in the immunology and metabolism datasets, HN is not. In addition, while HN is highly correlated with hit articles in applied physics, HC is not. Beyond variation across disciplines and the case for contextual analysis, we found that the categories demonstrating the highest percentage of hits are not robust with respect to varying parameters that define hit articles and the threshold for highly novel citation patterns.

@@

\section{Materials and Methods}
\emph{Bibliographic data} We have previously developed ERNIE, an open source knowledge platform into which we parse the Web of Science (WoS) Core Collection~\citep{Keserci371955}. WoS data stored in ERNIE spans the period 1900-2019 and consists of over 72 million publications. For this study, we generated an analytical dataset from years 1985 to 2005 from ERNIE. The total number of publications in this dataset was just over 25 million publications (25,134,073), which were then stratified by year of publication. For each of these years, we further restricted analysis to publications of type Article. Since WoS data also contains incomplete references or references that point at other indexes, we also considered only those references for which there were complete records~(Table \ref{tab:summary_data}). For example, WoS data for year 2005 contained 1,753,174 publications, which after restricting to type Article and considering only those references described above resulted in 916,573 publications, 6,095,594 unique references (set of references), and 17,167,347 total references (multiset of references). Given consistent trends in the data, Table \ref{tab:summary_data}, we analyzed the two boundary years (1985 and 2005) and the mid-point (1995) performing 1,000 simulations for each dataset. We also calculated the number of times each of these articles was cited in the first 8 years since publication.

% latex table generated in R 3.5.3 by xtable 1.8-4 package
% Fri May 17 14:28:31 2019
\begin{table}[ht]
\caption{Summary of base WoS Analytical Dataset.The number of publications of type Article (UP), unique references (UR), total references (TR) and the ratio of total references to unique references increases monotonically with each year indicating that both the number of documents and citation activity increase over time. Only publications of type Article with at least two references, and references with complete publication data were counted.} 
\label{tab:summary_data}
\centering
\begin{tabular}{|r rrrr|}
  \hline
year & UP & UR & TR & TR/UR \\ 
  \hline
1985 & 391,860 & 2,266,584 & 5,588,861 & 2.47 \\ 
  \hline
1986 & 402,309 & 2,316,451 & 5,708,796 & 2.46 \\ 
1987 & 412,936 & 2,427,347 & 5,998,513 & 2.47 \\ 
1988 & 426,001 & 2,545,647 & 6,354,917 & 2.50 \\ 
1989 & 443,144 & 2,673,092 & 6,749,319 & 2.52 \\ 
1990 & 458,768 & 2,827,517 & 7,209,413 & 2.55 \\ 
1991 & 477,712 & 2,977,784 & 7,729,776 & 2.60 \\ 
1992 & 492,181 & 3,134,109 & 8,188,940 & 2.61 \\ 
1993 & 504,488 & 3,278,102 & 8,676,583 & 2.65 \\ 
1994 & 523,660 & 3,458,072 & 9,255,748 & 2.68 \\ 
  \hline
1995 & 537,160 & 3,680,616 & 9,875,421 & 2.68 \\ 
  \hline
1996 & 663,110 & 4,144,581 & 11,641,286 & 2.81 \\ 
1997 & 677,077 & 4,340,733 & 12,135,104 & 2.80 \\ 
1998 & 693,531 & 4,573,584 & 12,728,629 & 2.78 \\ 
1999 & 709,827 & 4,784,024 & 13,280,828 & 2.78 \\ 
2000 & 721,926 & 5,008,842 & 13,810,746 & 2.76 \\ 
2001 & 727,816 & 5,203,078 & 14,261,189 & 2.74 \\ 
2002 & 747,287 & 5,464,045 & 15,001,390 & 2.75 \\ 
2003 & 786,284 & 5,773,756 & 16,024,652 & 2.78 \\ 
 2004 & 826,834 & 6,095,594 & 17,167,347 & 2.82 \\ 
   \hline
 2005 & 886,648 & 6,615,824 & 19,036,324 & 2.88 \\ 
 \hline
\end{tabular}
\end{table}

@@

\emph{Monte Carlo simulations, normalization of observed frequencies, annotations, and `hit' papers}. Analysis was performed on publications from 1985, 1995, and 2005. Building upon prior work~\cite{uzzi_atypical_2013}, all ${n \choose 2}$ reference pairs were generated for each publication, where $n$ is the number of cited references in the publication. These reference pairs were then mapped to the journals they were published in using ISSN numbers as identifiers. Where multiple ISSN numbers exist for a journal, the most frequently used one in the WoS was assigned to the journal. In addition, publications containing fewer than two references were discarded. Journal pair frequencies were summed across the dataset to create observed frequencies $(F_{obs})$. We developed a performant citation switching algorithm, \emph{runtime enhanced permuting citation switcher (repcs)} that randomly permuted citations grouped by year of publication to switch citations while preserving the year of publication for both articles and references, the number of publications and the number of references in each dataset and the disciplinary composition of the references in each dataset. We used the citation switching algorithm of ~\cite{uzzi_atypical_2013} referred to as \emph{umsj}~\citep{boyack_vs_uzzi_2014}, as a benchmark for the repcs algorithm using code kindly provided by the authors. A comparative analysis showed that while 10 simulations of the WoS 1985 dataset (391,860 publications) completed in 2,186 hours using the umsj algorithm, it completed in less than one hour using the repcs algorithm implemented on a Spark cluster. Using either the repcs or umsj algorithms for 10 simulations resulted in expected value coverage of 75\% of the observed journal pair frequencies. Subsequent rounds of development and testing resulted in 99\% coverage of observed journal pairs when 1,000 simulations were conducted. The runtime advantage was significant enough that we chose to use the repcs algorithm in our study  and generated expected values averaged from 1,000 simulations for improved coverage for every dataset we analyzed.

Averaging the result of 1,000 simulations for each dataset studied, z-scores were then calculated for each journal-pair using the formula $(F_{obs} - F_{exp})/\sigma$ where $F_{obs}$ is the observed frequency, $F_{exp}$ is the averaged simulated frequency, and $\sigma$ is the standard deviation of the simulated frequencies. As a result of these calculations, each publication becomes associated with a set of z-scores corresponding to the journal pairs derived from pairwise combinations of its cited references. Positional statistics of z-scores were calculated for each publication, which was then labeled according to conventionality and novelty: (i) HC if the median z-score exceeded the median of median z-scores for all publications and LC if the median z-score was equal to or less than the median of median z-scores for all publications, and (ii) HN if the tenth percentile of z-scores for a publication was less than zero, and LN if the tenth percentile of z-scores for a publication was greater than zero.

To consider the relationship between citation impact, conventionality, and novelty we calculated percentiles for the number of accumulated citations in the first 8 years since publication for each article we studied and stratified. We also investigated multiple definitions of hit articles with hits defined as the 1\%, 2\%, 5\%, and 10\% top-cited articles.  Also, mention different thresholds, 1st and 10th, for novelty threshold.

@@

\subsection{Model Misspecification and the Attributes of Disciplinary Context} We invoke model mis-specification arising from irrelevant references as a concern when disciplinary subsets are analyzed using a random graph model that generates expected co-citation frequencies from the full body of scientific literature. To address this issue, we first assembled three exemplar datasets using keyword searches for the terms `applied physics', `immunology', and `metabolism' (Materials and Methods). Using the WoS extended classificatation of 153 subject areas, we then analyzed the disciplinary composition of the references in these three datasets into the five research areas in the Web of Science (life sciences and biomedicine, physical sciences, social sciences, technology, and arts and humanities) for the years 1985, 1995, and 2005. We observed that 84.2\% of the references in the applied physics dataset were classified as belonging to the WoS physical sciences research area. Similarly, 93.7\% and 92.1\% of the references in the immunology and metabolism datasets belonged to the life sciences and biomedicine research area, indicating strongly that references are more likely belong to the same research area as the publications that cite them.

We  next analyzed the consequences of citation shuffling within a disciplinary set or all of the Web of Science.  References in publications belonging to these three datasets were classified using the WoS Extended Subjects classification of 153 subject areas and summarized as a frequency distribution. A single shuffle of the references in these datasets or all the references in the corresponding WoS year slice was performed, using either the rpcs or umsj algorithms, after which subject frequencies were computed again. The fold change in subject frequencies of references before and after shuffling was calculated for these groups using all 153 Subject categories and are summarized as boxplots in Fig \ref{fig:be}. For example, the applied physics dataset contained 1 reference labeled Genetics and Heredity . This proportion was preserved in the intra-network shuffle but increased 1496 times when the WoS background was used as a source of substitution during Monte Carlo simulation with the repcs algorithm. Similarly, with the metabolism dataset.which contained one reference labeled Philosophy that was increased to 661 after a single repcs shuffle. The data show convincingly that the disciplinary composition of references is preserved when citation shuffling is conducted within a disciplinary network but is significantly altered when the full WoS network is used as a source of substitution. A second inference is that the choice of algorithm is not relevant beyond runtime advantage but that the background network used for the shuffle is critical.

We predicted that model misspecification as measured by the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and simulated frequencies in a disciplinary network would be less than the divergence for observed and simulated frequencies in the WoS superset. The results in Table \ref{tab:kld} indicate that for the set of journals common to both a disciplinary network and the WoS superset, simulations under our model consistently have a lower K-L divergence, roughly half, compared to simulations that draw from the WoS superset (and its attendant substitutions that are ectopic with respect to field and discipline).

We used the criteria of Uzzi to calculate normalized journal pair frequencies (z-scores) and classify publications according to conventionality and novelty. We observe that z-score calculations are sensitive to the background network (disciplinary or WoS). Figure \ref{fig:scatter} shows that the z-scores for the same journal pair can be positive (negative) when computed with respect to one data set but be negative (positive) for another data set. The journal-pair z-scores in Figure \ref{fig:scatter} have consistent signs for both Immunology and WoS data sets in 71.4\% of the instances and different signs for 28.6\% of the journal pairs. When a journal-pair z-score is negative with respect to one dataset and positive with respect to another dataset, articles citing that pair are more likely to be deemed novel in the first instance and less likely to be deemed novel in the second instance.  It seems that any journal pair should either be indicative of novelty or not, but it should never have contradictory implications that depend on the reference dataset. We view, therefore, these contradictions as symptomatic of inappropriately sampling from a broad dataset and ignoring observed citation patterns.  Figure \ref{fig:scatter} reflects that the WoS data set has approximately 44,000 fewer negative z-scores than does the immunology data set, which contributes to its significantly lower percentage of high-novelty articles.  \emph{The plots I made of the cumulative z-score distributions for Immunology and WoS also could be used to demonstrate this phenomenon, although the scatter plot does this more clearly, and it is more striking.  The scatter plot also encodes more data, that is, the matched z-scores for each journal pair.}

Further, we stratified publications in these datasets by percentiles using the number of citations accumulated in the first 8 years since publication and according to conventionality and novelty in a second dimension Fig \ref{fig:Fig1} and note that [INSERT WHATEVER WE NOTE].

We addressed this consideration by analyzing disciplinary subsets of the scientific literature, thereby restricting random selection of references to only those references in the disciplinary network being studied.
\emph{What about z-scores?} 
We have conjectured that the approach of Uzzi et al. for generating journal-pair z-scores is misspecified in its sampling from a broad dataset and its disregard for the frequency with which journal pairs are citedOur principal consideration was to restrict model misspecification arising from disciplinarily irrelevant references. We addressed this consideration by analyzing disciplinary subsets of the scientific literature, thereby restricting random selection of references to only those references in the disciplinary network being studied.As observed in the Introduction, Monte Carlo simulations that use references from all publications do not account for observed citation practice (sentence needs to be made more eloquent)...

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=0.8\linewidth]{background-effect}     
\caption{\ Intra-network citation shuffling preserves the disciplinary composition of references. Publications of type Article belonging to the three disciplinary networks (i) Applied Physics (ii) Immunology (iii) Metabolism were subject to a single shuffle of all their cited references using either the cited references in these networks as a source of random substitutions (bg\_local) or references from all articles in WoS (bg\_allWoS). Citation shuffling was performed using either our algorithm (repcs) or that of Uzzi (umsj). The disciplinary composition of cited references before and after shuffling was measured as frequencies for each of 153 sub-disciplines (from the extended subject classification in WoS) and expressed as fold difference between citation counts grouped by subject for original (o) and shuffled (s) references using the formula (fold\_difference = $ifelse(o > s, o/s, s/o)$) and rounded to the nearest integer. A fold difference of 1 indicates that citation shuffling did not alter disciplinary composition. Data are shown for articles published in 1985. All eight boxplots are generated from 153 observations. Null values were set to 1. Note y-axis: $log_2$ scale.} 
\label{fig:be}
\end{figure}
